---
title: "GroupN_HM1"
author: "Michele Alessi, Gianmarco Alessio, Gabriele Codega, Leonardo Musini"
date: "2022-11-21"
output:
  html_document:
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## FSDS - Chapter 3

### Ex 3.12

Simulate a random sampling from a normal population distribution with several $n$ values to illustrate the law of large numbers.

**Solution**

We want to simulate a sample of $n=10$,$n=1000$,$n=1000000$ (the number of datasets to be simulated) and we generate an outcome ($\overline{y}$) with a normal population distribution of $mean=50$ and $var=10$. We took the mean of this outcomes for each 3 cases and to prove the law of large numbers we used *the convergence in probability* of $\overline{y}$ to $\mu$ for $n\to\infty$.

```{r}
set.seed(11)
mean(rnorm(10,50,10));
mean(rnorm(10^3, 50, 10));
mean(rnorm(10^7,50,10))
```

As we can see the mean converge to $\mu=50$ as $n$ increase.

## FSDS - Chapter 4

### Ex 4.48

For a simple random sample of $n$ subjects, explain why it is about $95\%$ likely that the sample proportion has error no more that $1/\sqrt{n}$ in estimating the population proportion. (*Hint*: To show this "$1/\sqrt{n}$ rule", find two standard errors when $\pi=0.50$, and explain how this compares to two standard errors at other values of $\pi$). Using this result, show that $n=1/M^2$ is a safe sample size for estimating a proportion to within $M$ with $95\%$ confidence.

**Solution**

The sampling distribution of the simple random sample $\hat{\pi}$ is approximately normal for large $n$. So the Pivotal Quantity

$$Z=\frac{\hat{\pi}-\pi}{\sqrt{\frac{\pi(1-\pi)}{n}}}$$

has an approximate standard normal distribution. Bounding this between standard normal percentiles, to construct a CI for $\pi$ after observing the simple random sample $\hat{\pi}$.

$$\hat{\pi} -\pi=\pm z_{\alpha/2}\sqrt{\frac{\pi(1-\pi)}{n}}$$
From here we use the simple approximate formula for large $n$ (Wald confidence interval) to easily compute the CI:

$$\hat{\pi}\pm z_{\alpha/2} \sqrt{\frac{\hat{\pi}(1-\hat{\pi})}{n}}$$
So from the exercise we assume the confidence level of $1-\alpha=0.95$ and since $95\%$ of the standard normal distribution falls between $-1.96$ and $1.96$ we have:

$$\hat{\pi}\pm 1.96 \sqrt{\frac{\hat{\pi}(1-\hat{\pi})}{n}}$$
We can observe $1.96\sim2.00$ (the exercise sugestion) and $\sqrt{\hat{\pi}(1-\hat{\pi})}\sim 0.50$ (considering $\hat{\pi}=0.50$) so:

$$\hat{\pi}\pm \frac{1}{\sqrt{n}}$$
And with a simple simulation we can verify that:

```{r}
set.seed(11)
simulation <- 100
size<-100
thresholdrule <- 1/sqrt(size)

standarderror <- function(phi){
y<-rbinom(simulation,size,phi)
p<-y/size
sd <- sqrt(p*(1-p)/size)
return (sd)
}

#standarderror(0.50)
#standarderror(0.10)


all(standarderror(0.50)<thresholdrule)
all(standarderror(0.10)<thresholdrule)

```

Then if we consider the sample size $n=1/M^2$ and at the same time we want $n$ to be able to estimate $\pi$. We have

$$\hat{\pi}\pm M$$


## FSDS - Chapter 5

### Ex 5.2
Refer to the previous exercise. Using uniform prior distributions, find the posterior probability 
that the population proportion believing in life after death is higher for females than for males.

**Solution**
Refer to the previous exercise we have:
1. 1017 of 1178 females said yes
1. 703 of 945 males said yes

So we can assume that $Y_1\sim binom(n_1,\pi_1)$ and $Y_2\sim binom(n_2,\pi_2)$,

```{r}
pi1=rbeta(10000000,1017,1178); #pi greco 1
pi2 =rbeta(10000000,703,945)
quantile(pi1-pi2,c(0.025,0.975))

mean(pi1 < pi2) #approximate posterior probability P(pi1<pi2)
mean(pi1 > pi2) #approximate posterior probability P(pi1>pi2)

```

So it plausible that $\pi_1>\pi_2$ with a $P(\pi_1>\pi_2)=0.99$





